---
title: "Lab10_McGuire"
author: "Alexandra McGuire"
date: "4/1/2020"
output: html_document
bibliography: data/ref_lab10.bib
---
### Methods
The following graphs were created using r, following tutorials and examples from various sources. The data was taken from Johns Hopkins University. All code, warnings, and messages have been excluded from this report. Graphs were created using both specific daily reports and time-series data for various counties, provinces, states, and countries.

Packages utilized include tidyverse, maps, mapdata, lubridate, viridis, wesanderson, plotly, gganimate, and transformr.

### Results
```{r setup, include=FALSE, warning=FALSE, message=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(maps)
library(mapdata)
library(lubridate)
library(viridis)
library(wesanderson)
library(plotly)
```

```{r, warning=FALSE, message=FALSE, echo=FALSE}
daily_report <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-31-2020.csv"))%>% 
  filter (Country_Region == "China") %>%
  rename(Long = "Long_") 
    
ggplot(daily_report, aes(x = Long, y = Lat, size = Confirmed/1000)) +
    borders("china", colour = "grey", fill = "grey90") +
    theme_bw() +
    geom_point(shape = 21, color='red', fill='red', alpha = 0.5) +
    labs(title = 'COVID-19 Confirmed cases in China',x = '', y = '',
        size="Cases (x1000)") +
    theme(legend.position = "right") +
    coord_fixed(ratio=1.5)
```
This graph demonstrates the confirmed COVID-19 cases in different China provinces as of March 31, 2020. 

```{r, warning=FALSE, message=FALSE, echo=FALSE}
daily_report <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-27-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Country_Region == "US") %>% 
  group_by(Province_State) %>% 
  summarize(Deaths = sum(Deaths)) %>% 
  mutate(Province_State = tolower(Province_State))

# load the US map data
us <- map_data("state")
# We need to join the us map data with our daily report to make one data frame/tibble
state_join <- left_join(us, daily_report, by = c("region" = "Province_State"))

# plot state map
ggplot(data = us, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  # Add data layer
  geom_polygon(data = state_join, aes(fill = Deaths), color = "white") +
  geom_polygon(color = "black", fill = NA) +
  scale_fill_gradient(trans = "log10") +
  labs(title = "COVID-19 Deaths in the US")
```
This map is colored according to the amount of deaths from COVID-19 in each US state as of March 27, 2020.

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(plotly)

daily_report <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-27-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Province_State %in% c("Massachusetts", "Rhode Island", "Vermont", "New Hampshire", "Maine", "Connecticut")) %>% 
  group_by(Admin2) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  mutate(Admin2 = tolower(Admin2))

us <- map_data("state")
NE_us <- subset(us, region %in% c("massachusetts", "rhode island", "vermont", "new hampshire", "maine", "connecticut"))
counties <- map_data("county")
NE_county <- subset(counties, region %in% c("massachusetts", "rhode island", "vermont", "new hampshire", "maine", "connecticut"))

state_join <- left_join(NE_county, daily_report, by = c("subregion" = "Admin2")) 

# plot state map
ggplot(data = NE_county, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  # Add data layer
  geom_polygon(data = state_join, aes(fill = Confirmed), color = "white") +
  scale_fill_gradientn(colours = 
                         wes_palette("Zissou1", 100, type = "continuous"),
                       trans = "log10") +
  labs(title = "COVID-19 Confirmed Cases in New England")
```
This map of New England is broken down into counties, and shaded according to the amount of confirmed COVID-19 cases within the county.

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(plotly)

ggplotly(
  ggplot(data = NE_county, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
# Add data layer
  geom_polygon(data = state_join, aes(fill = Confirmed), color = "black") +
    scale_fill_gradientn(colours = 
                         wes_palette("Zissou1", 100, type = "continuous")) +
  ggtitle("COVID-19 Cases in New England") +
# Cleaning up the graph
  labs(x=NULL, y=NULL) +
  theme(panel.border = element_blank()) +
  theme(panel.background = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(axis.text = element_blank())
)
```
This is an interactive version of the previous map, allowing for specific values of confirmed cases to be obtained by scrolling over the county of interest.

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(gganimate)
covid <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")) %>%
  rename(Province_State= "Province/State", Country_Region = "Country/Region") %>%
  filter (Country_Region == "China") %>%
           pivot_longer(-c(Province_State, Country_Region, Lat, Long),
                  names_to = "Date", values_to = "Confirmed") %>%
           mutate(Date = mdy(Date) - days(1),
                  Place = paste(Lat,Long,sep="_")) %>%
# Summarizes state and province information
             group_by(Place,Date) %>%
           summarise(cumulative_cases = ifelse(sum(Confirmed)>0,
                     sum(Confirmed),NA_real_),
                     Lat = mean(Lat),
                     Long = mean(Long)) %>%
           mutate(Pandemic_day = as.numeric(Date - min(Date)))

china_day <- ggplot(covid,aes(x = Long, y = Lat, size = cumulative_cases/1000)) +
  borders("china", colour = "gray50", fill = "grey90") +
  theme_bw() +
  geom_point(color='red', alpha = .5) +
  labs(title = 'Pandemic Day: {frame}',x = '', y = '',
       size="Cases (x1000)") +
  theme(legend.position = "right") +
  coord_fixed(ratio=1.5)+
  transition_time(Date) +
  enter_fade()
china_day
```
This map of China over time shows the levels of confirmed COVID-19 cases throughout the various provinces.

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(ggplot2)
library(gganimate)
library(transformr)
theme_set(theme_bw())

time_series_confirmed_long <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")) %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")  %>% 
               pivot_longer(-c(Province_State, Country_Region, Lat, Long),
                            names_to = "Date", values_to = "Confirmed") 

# Let's get the times series data for deaths

time_series_deaths_long <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv")) %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")  %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long),
               names_to = "Date", values_to = "Deaths")

time_series_recovered_long <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_recovered_global.csv")) %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region") %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long),
               names_to = "Date", values_to = "Recovered")

# Create Keys 

time_series_confirmed_long <- time_series_confirmed_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".", remove = FALSE)

time_series_deaths_long <- time_series_deaths_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".") %>% 
  select(Key, Deaths)

time_series_recovered_long <- time_series_recovered_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".") %>% 
  select(Key, Recovered)

# Join tables
time_series_long_joined <- full_join(time_series_confirmed_long,
              time_series_deaths_long, by = c("Key"))

time_series_long_joined <- full_join(time_series_long_joined,
              time_series_recovered_long, by = c("Key")) %>% 
    select(-Key)

# Reformat the data
time_series_long_joined$Date <- mdy(time_series_long_joined$Date)

# Create Report table with counts
time_series_long_joined_counts <- time_series_long_joined %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long, Date),
               names_to = "Report_Type", values_to = "Counts")

time_series_long_joined %>% 
    filter(Country_Region == "China") %>%
    group_by(Province_State,Date) %>% 
    summarise_at(c("Confirmed", "Deaths", "Recovered"), sum) %>% 
    ggplot(aes(x = Date,  y = Confirmed, color = Province_State)) + 
    geom_point() +
    geom_line() +
    ggtitle("Confirmed COVID-19 Cases in China by Province") +
    geom_point(aes(group = seq_along(Date))) +
    transition_reveal(Date) 
```
This line plot breaks down the increase in confirmed COVID-19 cases over time, and throughout provinces of China. Hubei noticably has the largest amount of cases, as Wuhan, the epicenter of the outbreak, is located within this region.

Application written in R [@RCoreTeam] using data from JHU [@JHU2020]. Also using Lab 10 Tutorial [@Blanchard2020], and packages R Markdown [@RMarkdownTeam], Ggplot2 [@Ggplot2Team], and R studio [@RStudioTeam].

# References

